Hydra Battles and Cie

### 2020-12-11 doc/part-hydras.tex
minor corrections

\begin{Coqsrc}
Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=

| battle_1 : forall i h  h', battle_r   B i  h h' -> 
                            battle B i h (S i) h'
| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
                                   battle B (S i) h'' j h'  ->
                                   battle B i h j h'.
\end{Coqsrc}

@@

\begin{itemize}
\item At each round, Hercules chops off the rightmost head of the hydra.
\item The battle is standard: at the round number $i$, the expected replication is $i$.
\end{itemize}

@@

But the  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length. However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
Thus, we can consider this development as a collaboration of proof with computation.

In the following lines, we try to show faithfully how we found the value of the number $N$.

The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.

\subsection{The beginning of Hostilities}
During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.

@@

A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
subtrees of height 1 or 0).
Thus, we introduce handy notations.

@@

\subsection{Computing \dots}
In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
of the current hydra.

@@

The following function returns the next configurarion of the game. 
Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.

@@

Compute test 3.
   (**
     = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
     : state
    *)

Compute test 4.
 (*
  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
     : state
 *)

Compute test 5.
 (*
   = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
     : state
 *)

Compute test 2000.
(*
  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
     : state
*)
\end{Coqsrc}

@@

 \begin{Coqsrc}
 Compute test 10.
(*

    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
     : state
*)
 \end{Coqsrc}

@@

\begin{Coqsrc}
Compute test 22.
(*

 = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
     : state
*)
\end{Coqsrc}

\begin{Coqsrc}
 Compute test 46.
(*

 = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
     : state

*)

Compute test 94.

(*

 = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
     : state

*) 
\end{Coqsrc}

@@

\begin{Coqsrc}
 Compute test 95.

(*

  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
     : state

*)
\end{Coqsrc}

@@

Compute test (doubleS 95).

(**
 = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
     : state
 *)

Compute test (iterate doubleS 2 95).

(*
  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
     : state
*)
\end{Coqsrc}

>>>>

\begin{Coqsrc}
Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
| battle_1 : forall i h  h', 
      battle_r   B i  h h' -> battle B i h (S i) h'
| battle_n : forall i h  j h' h'',  
      battle_r  B i h h''  ->
      battle B (S i) h'' j h'  ->
      battle B i h j h'.
\end{Coqsrc}

@@

\begin{itemize}
\item At each round, Hercules chops off the rightmost head of the hydra.
\item The battle is called \emph{standard}\footnote{This appellation is ours. If there is a better one, we will change it.}: at the round number $i$, the expected replication is $i$.
\end{itemize}

@@

But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length. However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
Thus, we can consider this development as a collaboration of proof with computation.
In the following lines, we show how we found experimentally the value of the number $N$.
The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.

\subsection{The Beginning of Hostilities}
During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.

@@

A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
subtrees of height 1 or 0).
Thus, we introduce a few handy notations.

@@

\subsection{Testing  \dots}
In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
of the current hydra.

@@

The following function returns the next configurarion of the game. 
Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
\texttt{step\_battle}, page~\pageref{lemma:step-battle}.

@@

Compute test 3.
\end{Coqsrc}

\begin{Coqanswer}
      = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
     : state 
\end{Coqanswer}

\begin{Coqsrc}
 Compute test 4.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 5.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 2000.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
     : state
\end{Coqanswer}

@@

 \begin{Coqsrc}
 Compute test 10.
\end{Coqsrc}

\begin{Coqanswer}
    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
     : state
 \end{Coqanswer}

@@

\begin{Coqsrc}
Compute test 22.
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
 Compute test 46.
\end{Coqsrc}

\begin{Coqanswer}
( = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 94.
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
     : state
\end{Coqanswer}

@@

\begin{Coqsrc}
 Compute test 95.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
     : state
\end{Coqanswer}

@@

Compute test (doubleS 95).
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test (iterate doubleS 2 95).
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
     : state
\end{Coqanswer}
